Mendelian randomization for causal inference in genome-wide association studies

### 2022-03-01 mr-review.Rmd
added text for grant_efficient_2020 regularization method

>>>>

@cheng_mr-ldp_2020 sought a probabilistic model that accounts for both linkage disequilibrium among SNP 
instruments and horizontal pleiotropy. They termed their method MR-LDP. Using an approach from @zhu_bayesian_2017, @cheng_mr-ldp_2020 derived 
an approximate likelihood for their regression models of the exposure and outcome on the SNP instruments. 
Incorporation of a random effect into their models accounted for the variance in causal effect estimates due 
to horizontal pleiotropy. They situate inference within an empirical Bayesian framework and present a 
parameter-expanded variational Bayes expectation-maximization algorithm for estimation [@liu1998parameter]. 
By jointly modeling the distribution of GWAS summary statistics and causal effects, @cheng_mr-ldp_2020 
efficiently accommodates multiple SNP instruments in linkage disequilibrium. One limitation of MR-LDP  
is its reliance on the InSIDE assumption and resulting inability to accommodate correlated horizontal pleiotropy.

@zhu_iterative_2021 present IMRP, a method for causal effect estimation and horizontal pleiotropy detection. 
They describe a test to distinguish vertical pleiotropy from horizontal pleiotropy. By combining this test 
with two-sample MR analysis of GWAS summary statistics, they develop an iterative procedure for estimating 
the causal effect of exposure on outcome and test for presence of horizontal pleiotropy in the SNP 
instruments. The iterative procedure alternates between 1) updating the estimate of the causal effect and 2) 
testing for horizontal pleiotropy with the included SNP instruments. At each test for horizontal pleiotropy, 
some SNP instruments may be discarded if they demonstrate evidence of horizontal pleiotropic effects. The 
procedure ends when there remain only SNP instruments that don't exhibit horizontal pleiotropy. In the authors' 
simulations and data analysis sections, their new method performs about as well as GSMR, IVW, and MR-PRESSO in the 
settings with and without balanced pleiotropy and with and without satisfaction of the InSIDE assumption [@zhu_iterative_2021]. Finally, the authors note that IMRP is three orders of magnitude faster than simulations-based MR-PRESSO.

@grant_efficient_2020 build on research from @lin2015regularization to present a method that 
applies a $\ell_1$ penalty to coefficients of covariates in two-sample MR with GWAS summary 
statistics. Importantly, their method doesn't penalize the SNP-exposure association. Instead, 
it shrinks towards zero those coefficients for covariates that demonstrate little pleiotropy. In so doing, 
they extend work from @burgess_multivariable_2015 and @burgess_re_2015 to get a causal effect estimator that needs no valid SNP instruments and robust to measured vertical pleiotropy and more statistically efficient
than previous MR methods.
